## Loading and preprocessing the data
The `activity.csv` file was loaded using `read.csv` function. The option of `header` was set to `TRUE` by default.
```{r}
setwd("C:/R_coursera/RepData_PeerAssessment1")
activity <- read.csv("./activity.csv")
```
## What is mean total number of steps taken per day?
A function of `sumnona` is constructed to calculate the sum of steps for each day ignoring the `NA`s in the data. The result is saved in `sumofsteps`.

```{r}
sumnona <- function(x) {sum(x, na.rm = TRUE)}
sumofsteps <- tapply(activity$steps, activity$date, sumnona)
sumofsteps
```
The histogram is constructed below using `breaks = 10`.
```{r}
hist(sumofsteps, breaks = 10, main = "Histogram of the total number of steps taken each day", xlab = "Total number of steps taken each day")
```

The mean of the total number of steps taken per day is:
```{r}
mean(sumofsteps)
```
The median of the total number of steps taken per day is:
```{r}
median(sumofsteps)
```
## What is the average daily activity pattern?
First, we need to make the `inteval` a factor variable.
```{r}
activity$interval <- factor(activity$interval)
```
Then we need to calculate the average of steps taken for each interval, averaged across all days. The result is saved in `meanofsteps`.
```{r}
meannona <- function(x) {mean(x, na.rm = TRUE)}
meanofsteps <- tapply(activity$steps, activity$interval, meannona)
```
To construct the a time series plot:
```{r}
plot(unlist(dimnames(meanofsteps)), meanofsteps, type = "l", xlab = "Interval (min)", ylab = "Average number of steps taken")
```

The maximum average number of steps occurs at:
```{r}
names(which.max(meanofsteps))
```
## Imputing missing values
The total number of missing values in the dataset:
```{r}
sum(is.na(activity$steps))
```
I chose to use the mean for that 5-minute interval to fill in all the `NA`s. First, copy all the data into a new dataset `newac`.
```{r}
newac <- activity
```
Then fill in all the `NA`s using the mean for that 5-minute interval.
```{r}
for (i in 1:nrow(newac)) {
      if (is.na(newac$steps[i])) {
            inte <- newac$interval[i]
            newval <- unname(meanofsteps[names(meanofsteps) == inte])
            newac$steps[i] <- newval
      }
}
```
Calulate the total number of steps taken each day. The results is save in `newsum`.
```{r}
newsum <- tapply(newac$steps, newac$date, sum)
newsum
```
The new histogram is constructed below using `breaks = 10`.
```{r}
hist(newsum, breaks = 10, main = "NEW histogram of the total number of steps taken each day", xlab = "Total number of steps taken each day")
```

By comparing this new histogram with the previous histogram, it is clear to see that the total number of steps taken each day is more normally distributed by replacing the `NA`s with the mean for that 5-minute interval.

The mean of total number of steps taken per day:
```{r}
tapply(newac$steps, newac$date, mean)
```
The median of total number of steps taken per day:
```{r}
tapply(newac$steps, newac$date, median)
```
## Are there differences in activity patterns between weekdays and weekends?
Identify which weekday each date is and save the result in the variable `whichday`.
```{r}
for (i in 1:nrow(newac)) {
      newac$whichday[i] <- weekdays(as.Date(newac$date[i]))
}
for (i in 1:nrow(newac)) {
      if (newac$whichday[i] == "Saturday" | newac$whichday[i] == "Sunday") {
            newac$whichday[i] <- "weekend"
      } else {
            newac$whichday[i] <- "weekday"
      }
}
```
Subset the new dataset into two sets: `wdays` and `wends`.
```{r}
wdays <- newac[newac$whichday == "weekday", ]
wends <- newac[newac$whichday == "weekend", ]
```
Construct the plots:
```{r}
par(mfrow = c(1, 2))
wdmean <- tapply(wdays$steps, wdays$interval, mean)
plot(unlist(dimnames(wdmean)), wdmean, type = "l", xlab = "Interval (min)", ylab = "Average number of steps taken", main = "Weekday")
wemean <- tapply(wends$steps, wends$interval, mean)
plot(unlist(dimnames(wemean)), wemean, type = "l", xlab = "Interval (min)", ylab = "Average number of steps taken", main = "Weekend")
```

By comparison, it shows more steps are taken during the weekdays and the number of steps in different time intervals is more evenly distributed during the day in the weekends than weekdays.
